Non-Monotone GLS Simulation: No \(A_{00}\)

### 2024-02-02 Notes/sim_no00.tex
Fixed gls_sim so that the covariance matrix reflects the SRS design. Tested to see if removing A_00 made estimator worse in sim_no00. It did.

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Understanding Semiparametric Constraints},
  pdfauthor={Caleb Leedy},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Understanding Semiparametric Constraints}
\author{Caleb Leedy}
\date{22 January 2024}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, sharp corners]}{\end{tcolorbox}}\fi

\hypertarget{sec-intro}{%
\section{Introduction}\label{sec-intro}}

The goal of this report is to understand how additional constraints
effect semiparametric models. Inspired by Dr.~Fuller's note, I set out
to see if we could assess the amount of efficiency loss due to using
semiparametric inference. First, let's define some notation.

\hypertarget{tbl-reftab}{}
\begin{longtable}[]{@{}llll@{}}
\caption{\label{tbl-reftab}This table shows the relationship between
different quantities of interest in each segment, their estimators, and
the corresponding coefficients of their estimators.}\tabularnewline
\toprule\noalign{}
Segment & Quantity of Interest & Estimator & Coefficient \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Segment & Quantity of Interest & Estimator & Coefficient \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(A_{00}\) & \(E[g \mid X]\) & \(\hat \gamma_{11}\) & \(c_{11}\) \\
\(A_{10}\) & \(E[g \mid X]\) & \(\hat \gamma_{21}\) & \(c_{21}\) \\
\(A_{10}\) & \(E[g \mid X, Y_1]\) & \(\hat \gamma_{22}\) & \(c_{22}\) \\
\(A_{01}\) & \(E[g \mid X]\) & \(\hat \gamma_{31}\) & \(c_{31}\) \\
\(A_{01}\) & \(E[g \mid X, Y_2]\) & \(\hat \gamma_{33}\) & \(c_{33}\) \\
\(A_{11}\) & \(E[g \mid X]\) & \(\hat \gamma_{41}\) & \(c_{41}\) \\
\(A_{11}\) & \(E[g \mid X, Y_1]\) & \(\hat \gamma_{42}\) & \(c_{42}\) \\
\(A_{11}\) & \(E[g \mid X, Y_2]\) & \(\hat \gamma_{43}\) & \(c_{43}\) \\
\(A_{11}\) & \(E[g \mid X, Y_1, Y_2]\) & \(\hat \gamma_{44}\) &
\(c_{44}\) \\
\end{longtable}

This is a similar framework to understanding the class of estimators
that we discussed in \verb|optest_update.pdf|, where we write the
estimator as a weighted average of different expectations with
functional coefficients.

\[\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta)E[g(Z)
\mid X] + \beta_1(\delta)E[g(Z) \mid X, Y_1] + \beta_2(\delta) E[g(Z)
\mid X, Y_2].\]

In this case, we assume that \(E[\beta_j(\delta)] = 0\) for
\(j = 0, 1, 2\).

However, now we write the estimator as

\begin{equation}\protect\hypertarget{eq-linest}{}{\hat \theta = \sum_{k,t} c_{kt}\hat \gamma_{kt}}\label{eq-linest}\end{equation}

where
\(\hat \gamma_{kt} = \frac{\delta_{ij}}{\pi_{ij}}E[g \mid G_{ij}(X, Y_1, Y_2)]\)
and \(ij\) corresponds to the segment \(A_{ij}\) associated with
\(\hat \gamma_{kt}\) in Table~\ref{tbl-reftab}. In general the estimator
has the following form,

\begin{align*}
\hat \theta = n^{-1} \sum_{i = 1}^n 
  &\left\{c_{11}\frac{\delta_{00}}{\pi_{00}} E[g \mid X]+ \right.\\
  &c_{21}\frac{\delta_{10}}{\pi_{10}} E[g \mid X]+ 
  c_{22}\frac{\delta_{10}}{\pi_{10}} E[g \mid X, Y_1]+ \\
  &c_{31}\frac{\delta_{01}}{\pi_{01}} E[g \mid X]+ 
  c_{33}\frac{\delta_{01}}{\pi_{01}} E[g \mid X, Y_2]+ \\
  &c_{41}\frac{\delta_{11}}{\pi_{11}} E[g \mid X]+ 
  c_{42}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_1]+
  c_{43}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_2]+
  \left.c_{44}\frac{\delta_{11}}{\pi_{11}} E[g \mid X, Y_1, Y_2]\right\}
\end{align*}

which we have condensed to Equation~\ref{eq-linest}.

\hypertarget{sec-methods}{%
\section{Methods}\label{sec-methods}}

As noted by Dr.~Fuller, we can understand parametric estimation as
solving the following constrained optimization problems:

\begin{align*}
  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
  \sum_{k,t} c_{kt} = 1 \\
       &\text{or} \\
  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
  \sum_{(k,t): (k,t) \neq (4, 4)} c_{kt} = 0 \text{ and } c_{44} = 1. \\
\end{align*}

To be robust to the outcome model, we can construct a similar
optimization problem with different constraints:

\begin{align*}
  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
  c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0,\\
       &\text{and } c_{44} = 1.
\end{align*}

Likewise to be robust to the response model, we can have the problem:

\begin{align*}
  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
  c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \\
       &\text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}.
\end{align*}

To be double robust (robust to the outcome and response model) we can
combine the last two constraints. This is summarized in
Table~\ref{tbl-mods}.

\hypertarget{tbl-mods}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2941}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7059}}@{}}
\caption{\label{tbl-mods}This table identifies the different constraints
for each model type.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Type
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Constraints
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Type
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Constraints
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Parametric 1 & \(\sum_{k, t} c_{kt} = 1\) \\
Parametric 2 &
\(\sum_{k, t: (k, t) \neq (4, 4)} c_{kt} = 0, c_{44} = 1\) \\
Outcome Robust &
\(c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, \text{ and } c_{44} = 1.\) \\
Response Robust &
\(c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}\) \\
Double Robust &
\(c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, c_{44} = 1, c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}\) \\
\end{longtable}

We can justify the constraints for the Outcome Robust model by rewriting
the estimator into the following form:

\begin{align*}
E_\delta[\hat \theta] &= n^{-1} \sum_{i = 1}^n E_\delta\left[
E[g \mid X] \left(\frac{\delta_{00}}{\pi_{00}} c_{11} +
\frac{\delta_{10}}{\pi_{10}}c_{21} \frac{\delta_{01}}{\pi_{01}}c_{31} +
\frac{\delta_{11}}{\pi_{11}}c_{41}\right)\right. + \\ 
&\qquad\qquad\qquad\qquad E[g \mid X, Y_1] \left(\frac{\delta_{10}}{\pi_{10}}c_{22} + 
\frac{\delta_{11}}{\pi_{11}}c_{42}\right) +\\ 
&\qquad\qquad\qquad\qquad E[g \mid X, Y_2] \left(\frac{\delta_{01}}{\pi_{01}}c_{33} + 
\frac{\delta_{11}}{\pi_{11}}c_{43}\right) +\\ 
&\qquad\qquad\qquad\qquad \left.g \frac{\delta_{11}}{\pi_{11}}c_{44}\right] \\
&= n^{-1} \sum_{i = 1}^n \left\{E[g \mid X](c_{11} + c_{21} + c_{31} + c_{41}) +
E[g \mid X, Y_1] (c_{22} + c_{42}) +
E[g \mid X, Y_2] (c_{33} + c_{43}) + g c_{44}\right\}
\end{align*}

Thus, if the Outcome Robust constraints are satisfied,
\(E_\delta[\hat \theta] = g\) when we take the expectation with respect
to the response variable \(\delta\). These constraints make sense
because if they are satisfied \(E[\hat \theta]\) does not depend on the
correct specification of the models for \(E[g \mid G_r(X, Y_1, Y_2)]\).

Likewise, we can justify the constraints for the Response Robust model
by rewriting Equation~\ref{eq-linest} into the form

\begin{align*}
E_Z [\hat \theta] &= n^{-1} \sum_{i = 1}^n E_Z\left[
\frac{\delta_{00}}{\pi_{00}}c_{11} E[g \mid X]\right. + \\
&\qquad\qquad\qquad\qquad\frac{\delta_{10}}{\pi_{10}}\left(c_{21}E[g \mid X] + 
  c_{22}E[g \mid X, Y_1]\right) +\\
&\qquad\qquad\qquad\qquad\frac{\delta_{01}}{\pi_{01}}\left(c_{31}E[g \mid X] + 
  c_{33}E[g \mid X, Y_2]\right) +\\
&\qquad\qquad\qquad\qquad\left.\frac{\delta_{11}}{\pi_{11}}\left(c_{41}E[g \mid X] + 
  c_{42}E[g \mid X, Y_1] + c_{43}E[g \mid X, Y_2] + c_{44}g\right)\right] \\
& n^{-1} \sum_{i = 1}^n \left(\frac{\delta_{00}}{\pi_{00}}c_{11} + 
\frac{\delta_{10}}{\pi_{10}}(c_{21} + c_{22}) + 
\frac{\delta_{01}}{\pi_{01}}(c_{31} + c_{33}) + 
\frac{\delta_{11}}{\pi_{11}}(c_{41} + c_{42} + c_{43} + c_{44})\right)g
\end{align*}

Thus, if the Response Robust constraints are satisfied,
\(E_X[\hat \theta] = g\) when we take the expectation with respect to
the outcome variable \(Z = (X, Y_1, Y_2)'\). These constraints make
sense because when they are satisfied \(E[\hat \theta]\) does not depend
on the correct response probabilities \(\pi = E[\delta]\).

\hypertarget{sec-simulations}{%
\section{Simulation Study}\label{sec-simulations}}

\hypertarget{simulation-1}{%
\subsection*{Simulation 1}\label{simulation-1}}
\addcontentsline{toc}{subsection}{Simulation 1}

@@

\begin{align*}
  \begin{bmatrix} x \\ e_1 \\ e_2 \end{bmatrix} 
  &\stackrel{ind}{\sim} N\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
  \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & \rho \\ 0 & \rho & 1 \end{bmatrix}\right) \\
  y_1 &= x + e_1 \\
  y_2 &= \theta + x + e_2 \\
\end{align*}

Furthermore, \(\pi_{11} = 0.2\) and \(\pi_{00} = 0.3, \pi_{10} = 0.4\),
and \(\pi_{01} = 0.1\). The goal of this simulation study is to find
\(\theta = E[Y_2]\). In other words, \(g(Z) = Y_2\). There are several
algorithms for comparison which are defined as the following:

\begin{align*}
  Oracle &= n^{-1} \sum_{i = 1}^n g(Z_i)\\
  CC &= \frac{\sum_{i = 1}^n \delta_{11} g(Z_i)}{\sum_{i = 1}^n \delta_{11}} \\
  IPW &= \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g(Z_i)\\
\end{align*}

Furthermore, we include four existing estimators that we have already
proposed in the past: WLS, Prop, PropInd, and SemiDelta.

The estimator WLS is a a weight least square estimator derived in the
following manner. We now consider a normal model:

\[ 
  \begin{pmatrix}
    x_i \\ e_{1i} \\ e_{2i}
  \end{pmatrix} \stackrel{ind}{\sim}
  N \left(
  \begin{bmatrix}
    0 \\ 0 \\ 0
  \end{bmatrix},
  \begin{bmatrix}
    1 & 0 & 0 \\
    0 & \sigma_{11} & \sigma_{12} \\ 
    0 & \sigma_{12} & \sigma_{22}
  \end{bmatrix},
  \right)
\]

and define \(y_{1i} = \theta_1 + x_i + e_{1i}\) and
\(y_{2i} = \theta_2 + x_i + e_{2i}\). Then \(b_1 = b_2 = 1\). We define
\(\bar z_k^{(ij)}\) as the mean of \(y_k\) in segment \(A_{ij}\). This
means that we have means \(\bar z_1^{(11)}\), \(\bar z_2^{(11)}\),
\(\bar z_1^{(10)}\), and \(\bar z_2^{(01)}\). Let
\(W = [\bar z_1^{(11)}, \bar z_2^{(11)}, \bar z_1^{(10)}, \bar z_2^{(01)}]'\),
then for \(n_{ij} = |A_{ij}|\), we have

\[Z - M \mu \sim N(\vec 0, V)\]

\[M = 
  \begin{bmatrix}
    1 & 0 \\
    0 & 1 \\
    1 & 0 \\
    0 & 1 \\
  \end{bmatrix}
  \text{ and }
  V = 
  \begin{bmatrix}
    \frac{\sigma_{11}}{n_{11}} & \frac{\sigma_{12}}{n_{11}} & 0 & 0 \\
    \frac{\sigma_{12}}{n_{11}} & \frac{\sigma_{22}}{n_{11}} & 0 & 0 \\
    0 & 0 & \frac{\sigma_{11}}{n_{10}} & 0 \\
    0 & 0 & 0 & \frac{\sigma_{22}}{n_{01}} \\
  \end{bmatrix}.
\]

Thus, the BLUE for \(\mu = [\mu_1, \mu_2]'\) is

\[\hat \mu = (M' V^{-1} M)^{-1} M' V^{-1} W.\]

Hence, WLS is \(\mu_2\) as \(g(X, Y_1, Y_2) = Y_2\).

The remaining three estimators are drived from the following expression
where the values for \(\beta\) are provided in Table~\ref{tbl-beta}.
These are good estimators to compare to because Prop is the original
proposed estimator. PropInd has the same form as Prop except the values
for \(\beta\) is different. PropInd shares the same values of \(\beta\)
as all of the new models with constraints. SemiDelta is useful because
it is the best estimator in general so far.

\[
\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + 
\beta_0(\delta, c_0)E[g(Z) \mid X] + 
\beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + 
\beta_2(\delta, c_2) E[g(Z)
\mid X, Y_2].
\]

\hypertarget{tbl-beta}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1698}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4151}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4151}}@{}}
\caption{\label{tbl-beta}This table displays the values of \(\beta\) for
different estimator types.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Estimator
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(\beta_0(\delta, c_0)\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(\beta_1(\delta, c_1)\)
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Estimator
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(\beta_0(\delta, c_0)\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(\beta_1(\delta, c_1)\)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Prop &
\(\left(1 - \frac{(\delta_{10} + \delta_{11})}{(\pi_{10} + \pi_{11})} - \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)\)
&
\(\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)\) \\
PropInd &
\(\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - \frac{(\delta_{01})}{(\pi_{01})} + \frac{\delta_{11}}{\pi_{11}}\right)\)
&
\(\left(\frac{\delta_{10}}{\pi_{10}} - \frac{\delta_{11}}{\pi_{11}}\right)\) \\
SemiDelta &
\(c_0\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{00}}{\pi_{00}}\right)\)
&
\(c_1\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{10}}{\pi_{10}}\right)\) \\
\end{longtable}

In the simulation results in Table~\ref{tbl-sim1}, the new results have
the same label as the value from the \texttt{Type} column in
Table~\ref{tbl-mods}.

\hypertarget{tbl-sim1}{}
\begin{longtable}[]{@{}lrrrr@{}}
\caption{\label{tbl-sim1}Results from Simulation 1. True values:
\(\theta = 5, \rho = 0.5\). The test conducted for the T-statistic and
P-value is a two sample test to see if the estimator is unbiased. This
simulation uses \(Y_1 = x + \varepsilon_1\),
\(Y_2 = \theta + x + \varepsilon_2\) where \(X \sim N(0, 1)\) and
\((\varepsilon_1, \varepsilon_2)\) come from a mean zero bivariate
normal distribution with unit variance and covariance \(\rho\). The
segments are unbalanced with: \(\pi_{11} = 0.2\), \(\pi_{10} = 0.4\),
\(\pi_{01} = 0.1\), and \(\pi_{00} = 0.3\).}\tabularnewline
\toprule\noalign{}
Algorithm & bias & sd & tstat & pval \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Algorithm & bias & sd & tstat & pval \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Oracle & -0.002 & 0.045 & -1.669 & 0.048 \\
CC & 0.001 & 0.083 & 0.263 & 0.396 \\
IPW & 0.010 & 0.357 & 0.872 & 0.192 \\
WLS & 0.000 & 0.054 & -0.065 & 0.474 \\
\(\hat \theta_{prop}\) & -0.002 & 0.063 & -0.768 & 0.221 \\
\(\hat \theta_{\delta}\) & -0.001 & 0.064 & -0.342 & 0.366 \\
Parametric 1 & -0.001 & 0.061 & -0.509 & 0.305 \\
Parametric 2 & -0.001 & 0.061 & -0.509 & 0.305 \\
Outcome Robust & -0.001 & 0.061 & -0.543 & 0.294 \\
Response Robust & -0.001 & 0.061 & -0.509 & 0.305 \\
Double Robust & -0.001 & 0.061 & -0.543 & 0.294 \\
\end{longtable}

\hypertarget{simulation-2}{%
\subsection*{Simulation 2}\label{simulation-2}}
\addcontentsline{toc}{subsection}{Simulation 2}

Next, we used the same simulation setup except that we are interested in
estimating \(\theta = E[g] = E[Y_1^2 Y_2]\). We do not use the WLS
estimator because it does not work.

>>>>

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}
\renewcommand{\arraystretch}{1.3}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Non-Monotone GLS Simulation: No A\_\{00\}},
  pdfauthor={Caleb Leedy},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Non-Monotone GLS Simulation: No \(A_{00}\)}
\author{Caleb Leedy}
\date{1 February 2024}

\begin{document}
\maketitle

\section{Introduction}\label{introduction}

The goal of this report is to conduct a simulation study to show the
validity of using a GLS estimator when intermediate models do not also
estimate the parameter in question. This setup is different from the
previous setup is a couple of ways:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  We do not use \(E[g \mid G_r(Z)]\) as the \(g\)-functions. Instead, we
  have \(g = (g_1,
  g_2, g_3)' = (X, Y_1, Y_2)'\).
\item
  We have fewer comparison estimators. Since we changed the
  \(g\)-functions it makes less sense to compare then to other
  estimators that are using different intermediate estimators.
\item
  We use a simple random sample (SRS) instead of a Poisson sample. For
  this setup, we have each segment being totally independent of each
  other. Each segment also has a fixed sample size of 250 instead of
  having segments with random sample sizes with a total observation
  count of 1000.
\item
  The GLS estimator is now only estimating \(\theta = E[Y_2] = E[g_3]\).
\end{enumerate}

\section{Notation and Setup}\label{notation-and-setup}

Let \(Z = (X, Y_1, Y_2)'\). We want to estimate the parameter
\(\theta = E[Y_2]\) where we may not always observe \(Y_1\) and \(Y_2\).
Define segments that contain observations in which the same variables
are observed as in Table~\ref{tbl-vars1}.

\begin{longtable}[]{@{}ll@{}}
\caption{This table identifies which variables are observed in each
segment. Since \(X\) is always observed, the subscript for each segment
identifies which of variables \(Y_1\) and \(Y_2\) are in the segment
based on the position of a 1.}\label{tbl-vars1}\tabularnewline
\toprule\noalign{}
Segment & Variables Observed \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Segment & Variables Observed \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(A_{00}\) & \(X\) \\
\(A_{10}\) & \(X, Y_1\) \\
\(A_{01}\) & \(X, Y_2\) \\
\(A_{11}\) & \(X, Y_1, Y_2\) \\
\end{longtable}

Let \(\delta_{i, j_1, j_2}\) be the sample inclusion indicator for
observation \(i\) in segment \(A_{j_1, j_2}\), and let
\(\pi_{j_1, j_2}\) be the probability of selecting an element into
\(A_{j_1, j_2}\).

We consider the vector \(g(Z) = (g_1(X), g_2(Y_1), g_3(Y_2))'\) and
\emph{for this simulation setup}, let \(g_1\), \(g_2\), and \(g_3\) all
be the identity function \(I()\). This means that
\(g(Z) = (X, Y_1, Y_2)'\). Notice, that we have
\(\theta = E[Y_2] = E[g_3]\). In each segment \(A_{j_1, j_2}\) we can
obtain estimators of some of the \(g\) elements. We have the following:

\begin{align*}
g_1^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_1(x_i) \\
g_2^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_2(y_{1i}) \\
g_3^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_3(y_{2i}) \\
g_1^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} g_1(x_i) \\
g_2^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} g_2(y_{1i}) \\
g_1^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} g_1(x_i) \\
g_3^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} g_2(y_{2i}) \\
g_1^{(00)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{00}}{\pi_{00}} g_1(x_i) \\
\end{align*}

This yields the following linear estimator,

\[ \hat g = Zg + e \]

\[\hat g = 
\begin{bmatrix}
g_1^{(11)} \\
g_2^{(11)} \\
g_3^{(11)} \\
g_1^{(10)} \\
g_2^{(10)} \\
g_1^{(01)} \\
g_3^{(01)} \\
g_1^{(00)} \\
\end{bmatrix},
Z = 
\begin{bmatrix}
1 & 0 & 0 \\
0 & 1 & 0 \\
0 & 0 & 1 \\
1 & 0 & 0 \\
0 & 1 & 0 \\
1 & 0 & 0 \\
0 & 0 & 1 \\
1 & 0 & 0 \\
\end{bmatrix},
E[e] = 0,
\text{ and } 
\Var(e) = n^{-1}
\begin{bmatrix}
V_{11} & 0 & 0 & 0 \\
0 & V_{10} & 0 & 0 \\
0 & 0 & V_{01} & 0 \\
0 & 0 & 0 & V_{00} \\
\end{bmatrix}.
\]

Here, we also have

\[
V_{11} = 
\begin{bmatrix}
\frac{1}{\pi_{11}} E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{11}} E[g_1g_2] -
E[g_1]E[g_2] & \frac{1}{\pi_{11}} E[g_1g_3] - E[g_1]E[g_3] \\
\frac{1}{\pi_{11}} E[g_1g_2] - E[g_1]E[g_2] & \frac{1}{\pi_{11}} E[g_2^2] -
E[g_2]^2 & \frac{1}{\pi_{11}} E[g_2g_3] - E[g_2]E[g_3] \\
\frac{1}{\pi_{11}} E[g_1g_3] - E[g_1]E[g_3] & \frac{1}{\pi_{11}} E[g_2g_3] -
E[g_2]E[g_3] & \frac{1}{\pi_{11}} E[g_3^2] - E[g_3]^2 \\
\end{bmatrix},
\] \[
V_{10} = 
\begin{bmatrix}
\frac{1}{\pi_{10}}E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{10}}E[g_1g_2] - E[g_1]E[g_2]\\
\frac{1}{\pi_{10}}E[g_1g_2] - E[g_1]E[g_2] & \frac{1}{\pi_{10}}E[g_2^2] - E[g_2]^2 \\
\end{bmatrix},
\] \[
V_{01} = 
\begin{bmatrix}
\frac{1}{\pi_{01}}E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{01}}E[g_1g_3] - E[g_1]E[g_3]\\
\frac{1}{\pi_{01}}E[g_1g_3] - E[g_1]E[g_3] & \frac{1}{\pi_{01}}E[g_3^2] - E[g_3]^2 \\
\end{bmatrix}, \text{ and }
V_{00} = 
\begin{bmatrix}
\frac{1}{\pi_{00}}E[g_1^2] - E[g_1]^2
\end{bmatrix}.
\]

\section{Simulation}\label{simulation}

@@

\begin{align*}
  \begin{bmatrix} x \\ e_1 \\ e_2 \end{bmatrix} 
  &\stackrel{ind}{\sim} N\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
  \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & \rho \\ 0 & \rho & 1 \end{bmatrix}\right) \\
  y_1 &= x + e_1 \\
  y_2 &= \mu + x + e_2 \\
\end{align*}

This yields outcome variables \(Y_1\) and \(Y_2\) that are correlated
both with \(X\) and additionally with each other. To generate the
missingness pattern, we select 250 observations into the four segments
independently.

\begin{longtable}[]{@{}lrrrr@{}}
\caption{Results from simulations study with independent equally sized
segments \(A_{11}\), \(A_{10}\), and \(A_{01}\) all of size \(n = 250\).
In this simulation we have the true mean of \(Y_2\) equal to \(\mu = 5\)
and the covariance between \(e_1\) and \(e_2\) is \(\rho = 0.5\). The
goal is to estimate \(E[Y_2] = \mu\). For the GLS estimation, we use the
true known covariance matrix (using the true values of \(\mu\) and
\(\rho\)), and we use g-functions \(g_1 = X\), \(g_2 = Y_1\) and
\(g_3 = Y_2\).}\tabularnewline
\toprule\noalign{}
Algorithm & Bias & SD & Tstat & Pval \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Algorithm & Bias & SD & Tstat & Pval \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Oracle & -0.001 & 0.049 & -0.818 & 0.207 \\
CC & -0.004 & 0.089 & -1.317 & 0.094 \\
IPW & -0.004 & 0.089 & -1.317 & 0.094 \\
GLS & -0.002 & 0.053 & -1.490 & 0.068 \\
\end{longtable}

A quick comparison between the standard deviation of the GLS estimator
and the standard deviation of the GLS estimator in \texttt{gls\_sim.qmd}
shows that this standard deviation is slightly larger, which means that
we do lose efficiency by not including \(A_{00}\).
